---
title: "My Project"
author: "Shelby Joji"
date: "November 1, 2018"
output: html_document
---

## R Markdown

The aim of this project is to develop a model that predicts the binary outcome: what factors predicts the presence of heart disease(CAD)?'

Source: https://www.drivendata.org/competitions/54/machine-learning-with-a-heart/page/109/

Project contains two datasets -  train and test. d1 contains all possible regressors for each unique patient ID. d2 contains outcome of all patient in d1.
We created train dataset by merging outcome(d2) with possible regressors(d1). Model is trained using train dataset. d3 is final test dataset to predict the outcome. Once we have merged, we do not need patient_id, which is the primary key for the dataset, as this is not a regressor or variable id


```{r}
library(readr)
d1 <- read_csv("train_values.csv")
d2 <- read_csv("train_labels.csv")
d3 <- read_csv("test_values.csv")
View(d1)
View(d2)

train<- total <- merge(d1,d2,by="patient_id")
train <- subset( train, select = -patient_id )

View(train)
#check data
str(train)
summary(train)
```

OBJECTIVE
Predict whether patient have heart disease (CAD) or not


_______________________________
Categorical variables
_______________________________

*Sex -  0 Female or  1 Male

thal - variable is character type

  This shows how well blood flows into your heart while exercising or at rest. Results of thallium stress test measuring blood flow to the heart, with possible 
        Possible values: normal, fixed_defect, reversible_defect
thal changed to thalach

*thalach
Since variable is of character type, we change it to 
          #-- Value 0: normal
          #-- Value 1: abnormal

```{r}
train$thalach[train$thal == "normal"] <- 0   
train$thalach[train$thal == "fixed_defect"] <- 1
train$thalach[train$thal == "reversible_defect"] <- 1
View(train)
train <- subset( train, select = -thal )

```


*chest_pain_type
        #-- Value 1: typical angina 
        #-- Value 2: atypical angina 
        #-- Value 3: non-anginal pain (if its duration is over 30 minutes or less than 5 seconds) 
        #-- Value 4: asymptomatic 

 
*resting_ekg_results
        #-- Value 0: normal 
        #-- Value 1: having ST-T wave abnormality (T wave inversions and/or ST elevation or depression of > 0.05 mV) 
        #-- Value 2: showing probable or definite left ventricular hypertrophy by Estes' criteria 

*slope_of_peak_exercise_st_segment 
        #-- Value 1: upsloping (does not signify anything)
        #-- Value 2: flat 
        #-- Value 3: downsloping (severe myocardial ischemia)
        
*exercise_induced_angina (1 = yes; 0 = no)

_______________________________
Continuous numeric variables
_______________________________

*serum_cholesterol_mg_per_dl - 126 (min) to 564(max) 

*oldpeak_eq_st_depression (ST depression induced by exercise relative to rest )
    - 0.00 (min)
    - 6.20 (max)

*max_heart_rate_achieved
    - 96 (min)
    - 202(max)

*Rest_BP - This should probably be systolic
    - 94 (min)
    - 180 (max)
*Age
    - 29 (min)
    - 77 (max)


_____________________________________
BINARY DEPENDENT variable
_____________________________________

*heart_disease_present - what we are trying to predict
    - 0 represents no heart disease present
    - 1 represents heart disease present
    



____________________________________________
EXPLORATORY ANALYSIS   PENDING
____________________________________________

____________________________________________
CHECK FOR MISSING VALUES      DONE
____________________________________________

Following code will generate a map for missing values. We use Amelia package for performing this analysis.

```{r}
library(Amelia)
sapply(train, function(x) sum(is.na(x)))

sapply(train, function(x) length(unique(x)))#unique values for all columns in dataset


missmap(train, main="Missing values vs observed")
```
______________________________________________
MODELING    PENDING
______________________________________________

We perform logistic regression due to the nature of binary dependent variable

Logistic regression
   - dependent variable is dichotomous 
   - explain relationship between response variable and one or more categorical(nominal), categorical(ordinal), interval or retio-level indpendent variables


ASSUMPTIONS
   - Response varibale is dichotomous  *MET*
   - No outliers in the data *MET*
   - There should be no multicolinearity (high correlation) among predictors .  Do a correlation matrix *TO BE CHECKED* This assuption is met as long as correlation coefficents are less than 0.90


```{r}
library(broom)

fullmodel<-glm(heart_disease_present ~ slope_of_peak_exercise_st_segment+thalach+resting_blood_pressure+chest_pain_type+num_major_vessels+fasting_blood_sugar_gt_120_mg_per_dl+resting_ekg_results+serum_cholesterol_mg_per_dl+oldpeak_eq_st_depression+sex+age+max_heart_rate_achieved+exercise_induced_angina+thalach,family = binomial(link = 'logit'), data=train)

tidy(fullmodel)
summary(fullmodel)


# to get odds ratio
require(MASS)
exp(cbind(Odd_Ratio = coef(fullmodel), confint(fullmodel)))

# create function to get probablity
logit2prob <- function(x){
  odds <- exp(x)
  prob <- odds / (1 + odds)
  return(prob)
}

# to get probablity
logit2prob(coef(fullmodel))

#confidence intervals
confint(fullmodel) # profile confidence intervals
confint.default(fullmodel)#Wald or Model-Based Confidence Intervals

exp(coef(fullmodel)) # exponentiated coefficients
exp(confint(fullmodel)) # 95% CI for exponentiated coefficients
predict(fullmodel, type="response") # predicted values
residuals(fullmodel, type="deviance") # residuals

#get odds ratio statistics
toOR <- function(object, ...) {
coef <- object$coef
se <- sqrt(diag(vcov(object)))
zscore <- coef / se
or <- exp(coef)
delta <- or * se
pvalue <- 2*pnorm(abs(zscore),lower.tail=FALSE)
loci <- coef - qnorm(.975) * se
upci <- coef + qnorm(.975) * se
ortab <- data.frame(or, delta, zscore, pvalue,
exp(loci), exp(upci))
round(ortab, 4)
}

#confidence interval shown is model based
toOR(fullmodel)



nothing <- glm(heart_disease_present ~ 1,family=binomial(link = 'logit'), data=train)
summary(nothing)

#backward regression
backwards = step(fullmodel) # Backwards selection is the default
summary(backwards)
formula(backwards)
back2<- glm(heart_disease_present ~ thalach + chest_pain_type + num_major_vessels  + oldpeak_eq_st_depression+ exercise_induced_angina,family=binomial(link = 'logit'), data=train)
summary(back2)

#forward regression
forwards = step(nothing,
scope=list(lower=formula(nothing),upper=formula(fullmodel)), direction="forward")
summary(forwards)



 bothways = step(nothing, list(lower=formula(nothing),upper=formula(fullmodel)),
direction="both",trace=0)
 summary(bothways)


formula(forwards)

formula(bothways)
  
both2<-glm(heart_disease_present ~ thalach + chest_pain_type + num_major_vessels  + oldpeak_eq_st_depression+ exercise_induced_angina,family=binomial(link = 'logit'), data=train)

summary(back2)
```

